---
title: "Missing Data"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r include=FALSE}
library(naniar)
library(tidyverse)

```

# Dealing with Missing Data 

Useful summary function from package 'naniar'
```{r}
miss_var_summary(airquality)
```


Replacing unusual values with missing values (NA) using mutate & ifelse
Replaces Ozone value <15 with NA
```{r}
head(airquality)

airquality2 <- airquality %>% 
  mutate(Ozone = ifelse(Ozone < 15, NA, Ozone))

head(airquality2)
```

##Count & Replace Strange Missing Values
Count missing values for -99, "N/A" and "n/a"
```{r}
#define messy dataframe
messy <- tibble::tribble(~x,  ~y,    ~z,
                          1,   "A",   -100,
                          3,   "N/A", -99,
                          NA,  NA,    -98,
                          -99, "E",   -101,
                          "n/a", "F",   -1)

 #count missing (in list) for each variable
miss_scan_count(messy,list(-99,"N/A","n/a"))
```

We can replace these inconsistent values with NA 
```{r}
messy %>% replace_with_na_at(
                   .vars = c("x","y","z"), 
                   ~.x %in% list(-99,"N/A","n/a"))

```

##Implicit missing values
Some missing values can be "hidden": rows that should be there are missing altogether
```{r}
hidden <- tibble::tribble(~player,~time,~score,
                          "jesse","morning",15,
                          "jesse",  "afternoon", 20,
                          "jesse","evening", 30,
                          "andy", " morning", 4,
                          "andy",  "afternoon", 100,
                          "nic",  "afternoon", 92)


# Use `complete()` on the `time` variable to make implicit missing values explicit
#adds NA for implicit missing
hidden %>% complete(player,time)
```
For the time variable we can see that there should be an entry for morning, afternoon, and evening for each player. However, we do not have complete information for andy (missing evening) or nic (missing morn/eve). We can replace these hidden missing values using complete()

More useful information about 'naniar' [here] (https://www.njtierney.com/post/2018/06/12/naniar-on-cran/)

